---
title: "Homework 2"
author: "Ashwini Varghese"
date: "2022-10-05"
output: github_document
editor_options: 
  chunk_output_type: console
---

## Prep:

We'll start by having a code chunk in the beginning that loads all the packages we will need for this homework.

```{r prep}
library (tidyverse)
```

## Problem 1:

We will start by reading in the datafile using the `readr` function from the `tidyverse` package and cleaning the data by using the `clean_names` function from the `janitor` package. We will also retain certain variables and convert the entry variable from a character variable to a logical variable. 

```{r subway}
subway = read_csv("./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv") 
    subway = janitor::clean_names(subway)
    col_types = cols(Route8 = "c", Route9 = "c", Route10 = "c", Route11 = "c")
    select(subway,
      line, station_name, station_latitude, station_longitude, 
      starts_with("route"), entry, exit_only, vending, entrance_type, ada)
    mutate(subway, entry = ifelse(entry == "YES", TRUE, FALSE))
subway
```






