---
title: "Homework 2"
author: "Ashwini Varghese"
date: "2022-10-05"
output: github_document
editor_options: 
  chunk_output_type: console
---

## Prep:

We'll start by having a code chunk in the beginning that loads all the packages we will need for this homework.

```{r prep}
library (tidyverse)
library(readxl)
```

## Problem 1:

We will start by reading in the datafile using the `readr` function from the `tidyverse` package and cleaning the data by using the `clean_names` function from the `janitor` package. We will also retain certain variables and convert the entry variable from a character variable to a logical variable. 

```{r subway}
subway = read_csv("./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv") %>% 
    janitor::clean_names() %>% 
    select( 
      line, station_name, station_latitude, station_longitude, 
      starts_with("route"), entry, exit_only, vending, entrance_type, ada) %>% 
    mutate(entry = ifelse(entry == "YES", TRUE, FALSE))
subway
```


## Problem 2:

Let's start by reading and cleaning the Mr. Trash Wheel and Professor Trash Wheel datasets. 

```{r Trashwheel}
trash = read_excel("./data/Trash_Wheel_Collection_Data.xlsx", sheet = "Mr. Trash Wheel", range = "A2:N549") %>% 
  janitor::clean_names() %>% 
  drop_na(dumpster) %>% 
  mutate(sports_balls = as.integer(round(sports_balls))) %>% 
  mutate(ID = "A")
```


```{r Professor}
professor = read_excel("./data/Trash_Wheel_Collection_Data.xlsx", sheet = "Professor Trash Wheel", range = "A2:M96") %>% 
  janitor::clean_names() %>% 
  drop_na(dumpster) %>% 
  mutate(ID = "B")
```

Next we will combine both datasets into one dataset.

```{r combo}
combo = merge(x = trash, y = professor, all = TRUE) %>% 
  select(ID, everything())
```

The new and combined dataset has 641 observations `nrow(combo)` and 15 variables `ncol(combo)`. All the variables exist both sets except for the _sports_balls_ variable; it came from the **trash** dataset. We can distinguish which observation is from which dataset by the _ID_ variable; an _ID_ value equal to A is for the **trash** dataset and an _ID_ value of B is for the **professor** dataset. 

To find the total weight of trash collected by Professor Trash Wheel, we can use the following code: `sum(subset(combo, ID == "B")$weight_tons)`, which gives us the sum of the _weight_tons_ variable restricted to the observations from the **Professor** dataset, identified by _ID = B_. The answer is 190.12 tons.

To find the total number of sports balls collected by Mr. Trash Wheel, we can use the following code: `sum(subset(combo, ID == "A")$sports_balls)`, which gives us the sum of the _sports_balls_ variable restricted to the observations from the **Trash** dataset, identified by _ID = A_. The answer is 6877 sports balls.


## Problem 3:

Clean up the pols-month file:

```{r pols}
pols = read_csv("./data/fivethirtyeight_datasets/pols-month.csv") %>% 
    janitor::clean_names() %>% 
    separate(mon, sep = "-", into = c("year", "month", "day")) %>%
    mutate(month = month.name[as.numeric(month)]) %>% 
    mutate(year = as.numeric(year)) %>% 
    mutate(month = factor(month, levels = month.name)) %>% 
    mutate(president = case_when(prez_gop == 1 ~ "gop", TRUE ~ "dem")) %>% 
    select(-day, -prez_gop, -prez_dem) %>% 
    arrange(year, month)
```

Clean up the snp file:

```{r snp}
snp = read_csv("./data/fivethirtyeight_datasets/snp.csv") %>% 
    janitor::clean_names() %>% 
    separate(date, sep = "/", into = c("month", "day", "year")) %>% 
    mutate(month = month.name[as.numeric(month)]) %>% 
    mutate(century = case_when(year < 16 ~ 2000, TRUE ~ 1900)) %>% 
    mutate(year = as.numeric(year)) %>% 
    mutate(year = year + century) %>% 
    select(year, month, close) %>% 
    mutate(month = factor(month, levels = month.name)) %>% 
    arrange(year, month) 
```

Now tidy the unemployment data:

```{r unemploy}
unemploy = read_csv("./data/fivethirtyeight_datasets/unemployment.csv") %>% 
    janitor::clean_names() %>% 
    pivot_longer(
      jan:dec,
      names_to = "month",
      values_to = "percent_unemploy") %>% 
    mutate(month = as.factor(month)) %>% 
    mutate(month = month.name[as.numeric(month)]) %>% 
    mutate(month = as.factor(month)) %>% 
    mutate(month = factor(month, levels = month.name)) %>% 
    arrange(year, month)
```







