---
title: "Homework 2"
author: "Ashwini Varghese"
date: "2022-10-05"
output: github_document
editor_options: 
  chunk_output_type: console
---

## Prep:

We'll start by having a code chunk in the beginning that loads all the packages we will need for this homework.

```{r prep}
library (tidyverse)
library(readxl)
```

## Problem 1:

We will start by reading in the datafile using the `readr` function from the `tidyverse` package and cleaning the data by using the `clean_names` function from the `janitor` package. We will also retain certain variables and convert the entry variable from a character variable to a logical variable. 

```{r subway}
subway = read_csv("./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv") %>% 
    janitor::clean_names() %>% 
    select( 
      line, station_name, station_latitude, station_longitude, 
      starts_with("route"), entry, exit_only, vending, entrance_type, ada) %>% 
    mutate(entry = ifelse(entry == "YES", TRUE, FALSE))
subway
```


## Problem 2:

Let's start by reading and cleaning the Mr. Trash Wheel and Professor Trash Wheel datasets. 

```{r Trashwheel}
trash = read_excel("./data/Trash_Wheel_Collection_Data.xlsx", sheet = "Mr. Trash Wheel", range = "A2:N549") %>% 
  janitor::clean_names() %>% 
  drop_na(dumpster) %>% 
  mutate(sports_balls = as.integer(round(sports_balls))) %>% 
  mutate(ID = "A")
```


```{r Professor}
professor = read_excel("./data/Trash_Wheel_Collection_Data.xlsx", sheet = "Professor Trash Wheel", range = "A2:M96") %>% 
  janitor::clean_names() %>% 
  drop_na(dumpster) %>% 
  mutate(ID = "B")
```





